---
title: "Analytics Presentation on Case Study for Employee Attrition"
author: "Presented by: Andrew Peng"
output: html_document
date: "2022-12-11"
---

```{r setup, include=FALSE}



# **Description**
# DDSAnalytics wants to leverage data science to conduct an analysis of exting employee data to build a model which can predict the Attrition. The Attrition model should achieves at least 60% Specificity and Sensitivity and it should also attain an RMSE less than $4000 for salaray.

```{r}
# Load Required Library
library("stringr")
library("dplyr")
library("ggplot2")
library("scales")
library("tidyverse")
library("tidyr")
library("plotly")
library("class")
library("caret")
library("e1071")
library("GGally")
library("Hmisc")
library("corrplot")
library("randomForest")
library("ROCR")
library("GGally")
library("ggcorrplot")
library("corrr")
library("knitr")
library("corrr")
library("tibble")
library("lattice")


## Importing the Data Set
```{r}
CaseStudy2 <- read.csv("C:/Users/andre/Desktop/CaseStudy2DDS/CaseStudy2-data.csv", header=TRUE)
No_Attrition <- read.csv("C:/Users/andre/Desktop/CaseStudy2DDS/CaseStudy2CompSet No Attrition.csv",header=TRUE)
No_Salary <- read.csv("C://Users/andre/Desktop/CaseStudy2DDS/CaseStudy2CompSet No Salary.csv",header=TRUE)

## Initial Data Analysis

dim(CaseStudy2)
head(CaseStudy2)
summary(CaseStudy2)
#Check for Missing Values
sapply(CaseStudy2, function(x) sum(is.na(x)))
#Check class type for each variable
sapply(CaseStudy2, class) 
#Check data frame dimensions and column data types
str(CaseStudy2)

#Changing Attrition to determine the retention rate
CaseStudy2$Attrition = as.factor(CaseStudy2$Attrition)
summary(CaseStudy2$Attrition)

CaseStudy2Job <- CaseStudy2[, -c(10,11,23,28)]



# Intial Analysis Summary:
# Data Structure: 870 oberservations (rows), 36 features (variables)
# Missing Data: there is no missing data 
# Data Type: there are two types of data exits in this dataset: character and integer
# Imbalance dataset: 730/870 = 87% employee stayed with the company and 140/870 = 16% employee made the decision to leave

# Attrition Plot

ggplot(data=CaseStudy2,aes(Attrition,after_stat(count),fill=Attrition))+ ggtitle("Attrition Histogram")+ geom_bar()


ggplot(CaseStudy2, aes(x=JobRole) + 
  theme_classic() +
  geom_bar(aes(fill=Attrition)) +
  ggtitle("Attrition Histogram by Job Role") +
  ylab("Number of Employees") +
  xlab("Job Role") +
  theme(axis.text.x = element_text(vjust = grid::unit(c(-4, -2, 0), "points")))


## Data Clean up and modification
# Remove zero variance predictors: EmployeeCount, EmployeeNumber, Over18, StandardHours #
CaseStudy2Clean <- CaseStudy2[, -c(10,11,23,28)]
No_Attrition <- No_Attrition[,-c(9,10,22,27)]
No_Salary <- No_Salary[,-c(10,11,22,27)]

# Chang data type
CaseStudy2Clean$Attrition <- ifelse(CaseStudy2Clean$Attrition == "Yes",1,0)


#Transform Categorical Variables to numerical values.

CaseStudy2Clean$Travel[CaseStudy2Clean$BusinessTravel== "Non-Travel"] <-1
CaseStudy2Clean$Travel[CaseStudy2Clean$BusinessTravel== "Travel_Rarely"] <-2
CaseStudy2Clean$Travel[CaseStudy2Clean$BusinessTravel== "Travel_Frequently"] <-3
CaseStudy2Clean$Depart[CaseStudy2Clean$Department== "Human Resources"] <-1
CaseStudy2Clean$Depart[CaseStudy2Clean$Department== "Research & Development"] <-2
CaseStudy2Clean$Depart[CaseStudy2Clean$Department== "Sales"] <-3
CaseStudy2Clean$Degree[CaseStudy2Clean$EducationField== "Human Resources"] <-1
CaseStudy2Clean$Degree[CaseStudy2Clean$EducationField== "Life Sciences"] <-2
CaseStudy2Clean$Degree[CaseStudy2Clean$EducationField== "Marketing"] <-3
CaseStudy2Clean$Degree[CaseStudy2Clean$EducationField== "Medical"] <-4
CaseStudy2Clean$Degree[CaseStudy2Clean$EducationField== "Other"] <-5
CaseStudy2Clean$Degree[CaseStudy2Clean$EducationField== "Technical Degree"] <-6
CaseStudy2$Clean$Sex[CaseStudy2Clean$Gender== "Female"] <-1
CaseStudy2$Clean$Sex[CaseStudy2Clean$Gender== "Male"] <-2
CaseStudy2Clean$Job[CaseStudy2Clean$JobRole== "Human Resources"] <-1
CaseStudy2Clean$Job[CaseStudy2Clean$JobRole== "Healthcare Representative"] <-2
CaseStudy2Clean$Job[CaseStudy2Clean$JobRole== "Laboratory Technician"] <-3
CaseStudy2Clean$Job[CaseStudy2Clean$JobRole== "Manager"] <-4
CaseStudy2Clean$Job[CaseStudy2Clean$JobRole== "Manufacturing Director"] <-5
CaseStudy2Clean$Job[CaseStudy2Clean$JobRole== "Research Director"] <-6
CaseStudy2Clean$Job[CaseStudy2Clean$JobRole== "Research Scientist"] <-7
CaseStudy2Clean$Job[CaseStudy2Clean$JobRole== "Sales Executive"] <-8
CaseStudy2Clean$Job[CaseStudy2Clean$JobRole== "Sales Representative"] <-9
CaseStudy2Clean$Martial[CaseStudy2Clean$MaritalStatus== "Divorced"] <-1
CaseStudy2Clean$Martial[CaseStudy2Clean$MaritalStatus== "Married"] <-2
CaseStudy2Clean$Martial[CaseStudy2Clean$MaritalStatus== "Single"] <-3
CaseStudy2Clean$OT[CaseStudy2Clean$OverTime== "Yes"] <-1
CaseStudy2Clean$OT[CaseStudy2Clean$OverTime== "No"] <-0

#Drop Transformed Variables
CaseStudy2drop <- CaseStudy2Clean %>% select(-one_of(c("ID","BusinessTravel", "Department", "EducationField", "Gender", "JobRole", "MaritalStatus","OverTime")))
str(CaseStudy2drop)

#Change variable "Depart" as numerical variable
CaseStudy2drop$Depart <- as.numeric(CaseStudy2drop$Depart)

#Create a heatmap 
CaseStudy2drop %>% 
  keep(is.numeric) %>%
  tidyr::drop_na() %>%  
  cor %>%
  corrplot(
    addCoef.col = "grey", number.digits = 2, number.cex = 0.7, method = "color",
    order = "hclust", title = "Variable Corr Heatmap", tl.srt = 50, tl.cex = 0.8, mar=c(0,0,1,0))


#Create a correlation matrix round them to 3 decimal places
correlationMatrix <- cor(CaseStudy2drop)
round(correlationMatrix, 3)

model=glm(Attrition~.,data = CaseStudy2drop, family= "binomial")
summary(model)

#Excludes Attribution correlation with itself and displays a summary of correlations to Attribution
correlationMatrix[2,2] = NA
summary(correlationMatrix[,2])


correlationMatrixframe=as.data.frame(correlationMatrix)

correlationMatrixframe %>% 
  keep(is.numeric) %>%
  tidyr::drop_na() %>%  
  cor %>%
  corrplot(
    addCoef.col = "grey", number.digits = 2, number.cex = 0.7, method = "color",
    order = "hclust", title = "Variable Corr Heatmap", tl.srt = 50, tl.cex = 0.8, mar=c(0,0,1,0))
 
#Create a Correlation Matrix with Significance Levels (p-value)

CaseStudy2pvalue <- rcorr(as.matrix(CaseStudy2drop))

CaseStudy2pvalue$r
summary(CaseStudy2pvalue$P)

CaseStudy2pvalue=as.data.frame(CaseStudy2pvalue)


# correlation between Attrition and all other variables to determine top 3 variables
CaseStudy2corr <- cor(CaseStudy2drop[-1], CaseStudy2drop$Attrition) 
print(CaseStudy2corr)

### Top 3 Factors Impacting Attrition are "OT(Overtime)", "Age", and "Material(Material Status)"



### Identify Job Specific Trends:


#Job Role vs Job Satisfaction 

CaseStudy2Job %>% ggplot() + geom_boxplot(aes(x = JobRole, y = JobSatisfaction , fill = JobRole)) +
 ggtitle("Job Role v Job Satisfaction") + ylab("Job Satisfaction ")

#Job Role vs Age

CaseStudy2Job %>% ggplot() + geom_boxplot(aes(x = JobRole, y = Age, fill = JobRole)) +
 ggtitle("Job Role v Age") + ylab("Age")

#Job Role vs Monthly Income
CaseStudy2Job %>% ggplot(aes(x = JobRole, y = MonthlyIncome)) +
  geom_boxplot() +
  labs(x = "Job Role", y = "Monthly Income",
       title = "Income Distribution by Job Role")




### Model Building and Prediction ###

# Variables
set.seed(5)
iter = 100
numks = 10
split = .70

# Create a matrix to hold the values from each run.
Accurary = matrix(nrow = iter, ncol = numks)
Specificity = matrix(nrow = iter, ncol = numks)
Sensitivity = matrix(nrow = iter, ncol = numks)

for(j in 1:iter)
{
  TrainingRows = sample(1:dim(CaseStudy2drop)[1],round(split * dim(CaseStudy2drop)[1])) # Calculate Training Rows
  trainset = CaseStudy2drop[TrainingRows,]  # Split into 2 seperate data frames. Include Training Rows
  testeset = CaseStudy2drop[-TrainingRows,]  # Exclude Training Rows (Testing Rows)
  for(i in 1:numks)
  {
    
    classifications = knn(trainset[,c(2,24,19,20,21,22)], testeset[,c(2,24,19,20,21,22)],
                          trainset$Attrition, prob = TRUE, k = i)
    
    table(testeset$Attrition, classifications)
    cm = confusionMatrix(table(testeset$Attrition, classifications))
    
    Accurary[j,i] = ((cm$table[1,1] + cm$table[2,2])) / ((cm$table[1,1] + cm$table[1,2]) + (cm$table[2,1] + cm$table[2,2]))
    Sensitivity[j,i] = cm$table[1,1] / (cm$table[1,1] + cm$table[2,1])
    Specificity[j,i] = cm$table[2,2] / (cm$table[1,2] + cm$table[2,2])
  }
  
}

AccuracyMean = colMeans(Accurary)
SpecMean = colMeans(Specificity)
SensMean = colMeans(Sensitivity)
plot(seq(1,numks,1),AccuracyMean, type = "l")
which.max(AccuracyMean)
max(AccuracyMean)
plot(seq(1,numks,1),SpecMean, type = "l")
which.max(SpecMean)
max(SpecMean)
plot(seq(1,numks,1),SensMean, type = "l")
which.max(SensMean)
max(SensMean)


## Naive Bayes Model
```{r Naive Bayes, echo=FALSE, message=FALSE, warning=FALSE}
# Variables
set.seed(5)
iter = 100
split = .70
# Create a matrix to hold the values from each run.
AccuVect = vector(length = iter)
SensVect = vector(length = iter)
SpecVect = vector(length = iter)


for(j in 1:iter)
{
  
CaseStudy2drop = mutate_if(CaseStudy2drop, is.numeric, as.factor)
CaseStudy2drop = mutate_if(CaseStudy2drop, is.integer, as.factor)

  
  TrainingRows = sample(1:dim(CaseStudy2drop)[1],round(split * dim(CaseStudy2drop)[1])) # Calculate Training Rows
  Navtrain = CaseStudy2drop[TrainingRows,]  # Split into 2 seperate data frames. Include Training Rows
  Navtest = CaseStudy2drop[-TrainingRows,]  # Exclude Training Rows (Testing Rows)
  
  nbm <- naiveBayes(Navtrain[,c(2,24,19,22)],Navtrain$Attrition)
  
  # Predict outcomes for Testing data set.
  Navtest$predict_outcome = predict(nbm,Navtest)
  
  classifications = predict(nbm,Navtest)
  # classifications
  table(Navtest$Attrition, classifications)
  cm = confusionMatrix(table(Navtest$Attrition, classifications))
  
  AccuVect[j] = ((cm$table[1,1] + cm$table[2,2])) / ((cm$table[1,1] + cm$table[1,2]) + (cm$table[2,1] + cm$table[2,2]))
  SensVect[j] = cm$table[1,1] / (cm$table[1,1] + cm$table[2,1])
  SpecVect[j] = cm$table[2,2] / (cm$table[1,2] + cm$table[2,2])
}

mean(AccuVect)
mean(SpecVect)
mean(SensVect)
plot(seq(1,length(AccuVect),1),AccuVect, type = "l", main = "Accuracy From NB Model", xlab = "Iteration", ylab = "Accuracy")
plot(seq(1,length(SpecVect),1),SpecVect, type = "l", main = "Specificity From NB Model", xlab = "Iteration", ylab = "Specificity")
plot(seq(1,length(SensVect),1),SensVect, type = "l", main = "Sensitivity From NB Model", xlab = "Iteration", ylab = "Sensitivity")

#Create predictions and write to CSV file


# Transform date with derived features to run predict model

No_Attrition$Travel[No_Attrition$BusinessTravel== "Non-Travel"] <-1
No_Attrition$Travel[No_Attrition$BusinessTravel== "Travel_Rarely"] <-2
No_Attrition$Travel[No_Attrition$BusinessTravel== "Travel_Frequently"] <-3
No_Attrition$Depart[No_Attrition$Department== "Human Resources"] <-1
No_Attrition$Depart[No_Attrition$Department== "Research & Development"] <-2
No_Attrition$Depart[No_Attrition$Department== "Sales"] <-3
No_Attrition$Degree[No_Attrition$EducationField== "Human Resources"] <-1
No_Attrition$Degree[No_Attrition$EducationField== "Life Sciences"] <-2
No_Attrition$Degree[No_Attrition$EducationField== "Marketing"] <-3
No_Attrition$Degree[No_Attrition$EducationField== "Medical"] <-4
No_Attrition$Degree[No_Attrition$EducationField== "Other"] <-5
No_Attrition$Degree[No_Attrition$EducationField== "Technical Degree"] <-6
No_Attrition$Sex[No_Attrition$Gender== "Female"] <-1
No_Attrition$Sex[No_Attrition$Gender== "Male"] <-2
No_Attrition$Job[No_Attrition$JobRole== "Human Resources"] <-1
No_Attrition$Job[No_Attrition$JobRole== "Healthcare Representative"] <-2
No_Attrition$Job[No_Attrition$JobRole== "Laboratory Technician"] <-3
No_Attrition$Job[No_Attrition$JobRole== "Manager"] <-4
No_Attrition$Job[No_Attrition$JobRole== "Manufacturing Director"] <-5
No_Attrition$Job[No_Attrition$JobRole== "Research Director"] <-6
No_Attrition$Job[No_Attrition$JobRole== "Research Scientist"] <-7
No_Attrition$Job[No_Attrition$JobRole== "Sales Executive"] <-8
No_Attrition$Job[No_Attrition$JobRole== "Sales Representative"] <-9
No_Attrition$Martial[No_Attrition$MaritalStatus== "Divorced"] <-1
No_Attrition$Martial[No_Attrition$MaritalStatus== "Married"] <-2
No_Attrition$Martial[No_Attrition$MaritalStatus== "Single"] <-3
No_Attrition$OT[No_Attrition$OverTime== "Yes"] <-1
No_Attrition$OT[No_Attrition$OverTime== "No"] <-0

#Drop Transformed Variables
No_Attrition <- No_Attrition %>% select(-one_of(c("ID","BusinessTravel", "Department", "EducationField", "Gender", "JobRole", "MaritalStatus","OverTime")))
str(No_Attrition)

#New Model to predict Attrition - Export CSV FILE#
```{r}
No_Attrition['Attrition'] <- as.factor(NA)

PredictAtt = predict(classifications, No_Attrition)
PredictAtt <- cbind(No_Attrition$ID, PredictAtt)
colnames(PredictAtt) <- c("Model", "Attrition")
head(PredictAtt)
write.csv(PredictAtt , "C:/Users/andre/Desktop/CaseStudy2DDS/Case2PredictionsPeng Attrition.csv", row.names = FALSE)
```


```{r}
## Creating a model for Salary

CaseStudy2 %>% ggplot(aes(y = MonthlyIncome, x = JobLevel)) + geom_point() + geom_smooth()
CaseStudy2 %>% ggplot(aes(y = MonthlyIncome, x = JobRole)) + geom_boxplot()
CaseStudy2 %>% ggplot(aes(y = MonthlyIncome, x = YearsAtCompany)) + geom_point() + geom_smooth()
CaseStudy2 %>% ggplot(aes(y = MonthlyIncome, x = MonthlyIncome)) + geom_point() + geom_smooth()
CaseStudy2 %>% ggplot(aes(y = MonthlyIncome, x = YearsInCurrentRole)) + geom_point() + geom_smooth()

set.seed(3)
iter = 100
# Create a matrix to hold the values from each run.
ObserVect = vector(length = iter)
PredVect = vector(length = iter)
for(j in 1:iter)
{
  TrainingRows = sample(1:dim(CaseStudy2)[1],dim(CaseStudy2)[1]-1) # Calculate Training Rows (leave one out)
  SalTrain = CaseStudy2[TrainingRows,]  # Split into 2 seperate data frames. Include Training Rows
  SalTest = CaseStudy2[-TrainingRows,]  # Exclude Training Rows (Testing Rows)
  
  MI_fit <- lm(MonthlyIncome~JobLevel+JobRole+YearsAtCompany+MonthlyIncome+YearsInCurrentRole, data = SalTrain)
  ObserVect[j] <- SalTest$MonthlyIncome
  PredVect[j] <- predict(MI_fit, newdata = SalTest)
  
}


mi_model_df <- data.frame(ObserVect, PredVect)
# Calculate Resisduals
mi_model_df$Res <- mi_model_df$ObserVect - mi_model_df$PredVect
# Square the Residuals
mi_model_df$ResSQ = mi_model_df$Res^2
# Calculate the RMSE
sqrt(mean(mi_model_df$ResSQ))
summary(MI_fit)
## RMSE = 1145.198
#Confidence Intervals
confint(MI_fit)

Salary_Pred <- predict(MI_fit, newdata = No_Salary)

SalaryMerg <- merge(No_Salary$ID, Salary_Pred, by=0, all=TRUE) 
#dim(SalaryMerg)
#head(SalaryMerg,2)
# renaming the merged variable names
colnames(SalaryMerg)[colnames(SalaryMerg)=="Row.names"] <- "RowNumber"
colnames(SalaryMerg)[colnames(SalaryMerg)=="x"] <- "ID"
colnames(SalaryMerg)[colnames(SalaryMerg)=="y"] <- "MonthlyIncome"
SalaryMerg <- SalaryMerg[order(SalaryMerg$ID),] %>% dplyr::select(ID, MonthlyIncome)
#dim(SalaryMerg)
#head(SalaryMerg,2)


write.csv(SalaryMerg, file = "C:/Users/andre/Desktop/CaseStudy2DDS/Case2PredictionsPeng Salary.csv", row.names = FALSE)